### 2016-10-09 text/answers_to_reviewers.Rmd

<!-- ```{r, results="hide", warnings=FALSE, echo=FALSE, message=FALSE}
source('../R/functions-analyses.R')
cleanbib()
myRefs   <-  listBibs()
exporBibs(myRefs, erase=FALSE)
```
 -->
August 16, 2016

### 2016-10-11 text/MS.Rmd
added github doi to text + few typos

*1. Animal collections*  
To collect our study organisms, roughened acetate sheets secured to the underside of 6 mm thick PVC sheets suspended from floating pontoons at a depth of 1 m were deployed at Manly Boat Harbour, Queensland, Australia (27˚27' S, 153˚11' E) and checked regularly to remove fouling organisms other than our species of interest: *Bugula neritina* and *B. stolonifera* (arborescent bryozoans that have a three dimensional tree-like growth form), *Hippopodina iririkiensis* (an encrusting bryozoan that grows mostly in two dimensions as flat disc), and an unidentified species of sponge belonging to the Microcionidae family (`r myCite('hart2009ecology')`; again a flat species that grows mostly in two dimensions) (Table 1). After 6 weeks in the field (average SST 22 ˚C), sheets bearing colonies were returned to the University of Queensland where they were kept in aerated seawater tanks (18 ˚C) for up to 48 h prior to measurements. For encrusting colonies, sections of acetate bearing whole colonies were then cut from the sheets and fragmented (see e.g. `r myCite(c('nakaya2005procb', 'white2011amnat'))` for details and tests of potential experimental artifacts). Arborescent colonies were cut directly from the acetate sheets and fragmented. We focused on non-reproductive colonies of bryozoans (at that time of year 6 weeks was not sufficient time for colonies to show reproduction) to avoid the potentially complicating effects that reproduction might bring. For the sponge, we found no evidence of reproduction, though we cannot rule out sperm production.

@@

*3. Respiration measurements*  
Individual metabolic rates (converted to $\mu$l $\textrm{O}_\textrm{2}$ h^-1^, *n* = `r nrow(numbersList$metRates)`), $B_i$, were measured as rate of oxygen consumption, $\dot{V}\textrm{O}_\textrm{2}$, of colony fragments at nominal temperatures of 10 or 25 ˚C. All individuals (regardless of size) and species were randomly assigned to test temperatures, and we used replicate vials and controlled temperature cabinets (see below). Measurements were conducted using a 144-channel PreSens Sensor Dish Reader (SDR, AS-1 Scientific, Wellington, New Zealand) according to standard techniques `r myCite(c('koster2008meps', 'lighton2008book', 'white2011amnat'))`.  $\dot{V}\textrm{O}_\textrm{2}$ was measured by placing the colony fragment in a glass vial containing 0.2$\mu$m filtered sea water and a non-consumptive $\textrm{O}_\textrm{2}$ sensor spot and calculated from the rate of change of $\textrm{O}_\textrm{2}$ saturation ($m_a$, \% h^‑1^) as:

@@

where $m_b$ is the rate of change of $\textrm{O}_\textrm{2}$ saturation for blank vials containing no animals (\% h^-1^), ${\beta}_{\textrm{O}_\textrm{2}}$ is the oxygen capacitance of air-saturated sea water at 10 ˚C (6.4 ml l^-1^) or 25 ˚C (4.74 ml l^-1^) `r myCite('cameron1986book')` and $V$ is water volume (vials were 5 ml in volume and water volume was calculated by subtracting the volume of acetate and animals). At least four blank vials were recorded simultaneously with each 24-well plate to account for microbial oxygen consumption and sensor spots were calibrated with air-saturated seawater (100\% A.S.) and water containing 2\% sodium sulfite (0\% A.S.). After a 1 h settling-in period (i.e. time necessary for the system to come into equilibrium), $\dot{V}\textrm{O}_\textrm{2}$ measurements lasted up to 20 h depending on individual mass (tiny individuals require much longer measurement periods in order to detect $\dot{V}\textrm{O}_\textrm{2}$). Following measurement of $\dot{V}\textrm{O}_\textrm{2}$, colony fragments were blotted dry and weighed to 0.1 mg (Sartorius A 200 S, Sartorius AG, Göttingen, Germany). All measurements of $\dot{V}\textrm{O}_\textrm{2}$ were made in one of four dark constant temperature cabinets, with temperatures within a run randomized among cabinets.

@@

where mass, $\textrm{ln}M_i$ (mg), is a continuous variable, while temperature (two levels: 10 ˚C and 25 ˚C) and species (four levels corresponding to the four species) are categorical variables. With this model, we allow each species to have a different mass-scaling slope that may change with temperature; we also account for species- and temperature-specific deviations on the metabolic normalization (i.e. model intercept). We also included date of experiment, $\Delta D$ as a random effect on the metabolic normalization to help control for natural experimental variability. We then performed a model selection procedure (see below) by comparing the above full model with a series of nested two- and one-way interaction models in order to formally test for mass-temperature interactions in metabolic rates among different species.

All models were initially fitted using maximum likelihood `r myCite('zuur2009book')` with the R package *lme4* version 1.1-8 `r myCite('lme4')`. Each nested model was compared against the full model using likelihood ratio tests, and models were considered significantly different if *p* < 0.05. We then fitted the best model in a Bayesian framework by calling *JAGS* version 3.4.0 from the R package *R2jags* version 0.05-03 `r myCite('r2jags')` in order to derive posterior distributions and associated 95\% credible intervals (CIs) for the fitted parameters. Random effects were assumed to be normally distributed, with means of 0. Fitted parameters were assigned priors that were vague (i.e. locally uniform over the region supported by the likelihood) `r myCite('kruschke2014book')`. The posterior distributions of model parameters were estimated using Markov chain Monte Carlo (MCMC) methods by constructing three chains of 500,000 steps each, including 250,000-step burn-in periods. Chains were then updated with 500,000 steps each until convergence was reached using 250-thinning step interval, so a total of 6,000 steps were retained to estimate posterior distributions (i.e. 3 $\times$ 500,000/250 = 6,000). All data and code necessary to reproduce this paper, its analyses, tables and figures can be obtained on GitHub https://github.com/dbarneche/MTRBrEs.

@@

The model selection indicated that the three-way interaction model (equation 2) performed significantly better than all nested two- and one-way interaction models (Table 2), thus implying that different species present different interactions between mass-scaling and temperature dependence. Particularly, the best model yields no interaction between mass scaling and temperature dependence for *Bugula stolonifera* and the Microcionidae sponge (large overlap between 95\% credible intervals for the mass-scaling slopes, $\alpha$) (Fig. 2). Still, while the mass-scaling slopes are statistically the same between temperatures for these species, there is evidence for higher metabolic normalizations (i.e. intercepts), $B_o$, at higher temperatures (though marginally significant for Microcionidae, i.e. slight overlap between 95\% credible intervals; bottom-right quadrants in Fig. 2). For example, metabolic rates increase `r rounded(exp(averageEstimates$lnB25[averageEstimates$species == 'Stolonifera'])/exp(averageEstimates$lnB10[averageEstimates$species == 'Stolonifera']))`-fold between 10 ˚C and 25 ˚C for *B. stolonifera*, yielding a $Q_{10}$ of `r rounded(q10andErs$q10[q10andErs$species == 'Stolonifera'])`, which is equivalent to an activation energy of `r rounded(q10andErs$er[q10andErs$species == 'Stolonifera'], 2)` eV `r myCite(c('gillooly2001','yvondurocher2012a'))`. The best model indicates that the difference in metabolic rates between 25 ˚C and 10 ˚C for *Hippopodina iririkiensis* decreases, on average, `r rounded(tempDeltas[1,'tempDelta'])`-fold moving from `r round(exp(2))` to `r round(exp(5))` mg. For *Bugula neritina*, however, rates increase `r rounded(1/tempDeltas[3,'tempDelta'])`-fold along the same body-mass gradient.

Average posterior estimates of mass-scaling exponents (top-left quadrants in Fig. 2) are inconsistent among the studied species and between growth forms (2D and 3D). For example, for organisms that exhibit 2D growth, *Hippopodina iririkiensis* shows exponents that encompass both 0.67 (Fractal geometry model prediction for 2D organisms; `r myCite('koontz2009amjbot')`) and 0.75 (Fractal geometry model prediction for 3D organisms; `r myCite(c('west1997science', 'west1999science'))`) at 10 ˚C (95\% CI: `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's10_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's10_97.5'], 2)`), but that are significantly shallower than 0.75 at 25 ˚C (95\% CI: `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's25_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's25_97.5'], 2)`). Microcionidae presents exponents that only overlap 0.67 at 10 ˚C, but not at 25 ˚C (95\% CI at 10 ˚C: `r rounded(ci95Estimates[ci95Estimates$species == 'Sponge', 's10_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Sponge', 's10_97.5'], 2)`; 95\% CI at 25 ˚C: `r rounded(ci95Estimates[ci95Estimates$species == 'Sponge', 's25_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Sponge', 's25_97.5'], 2)`). On the other hand, for 3D organisms, *Bugula neritina* presents scaling exponents that overlap both 0.67 and 0.75 at 25 ˚C (95\% CI: `r rounded(ci95Estimates[ci95Estimates$species == 'Bugula', 's25_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Bugula', 's25_97.5'], 2)`), but are significantly shallower at 10 ˚C (95\% CI: `r rounded(ci95Estimates[ci95Estimates$species == 'Bugula', 's10_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Bugula', 's10_97.5'], 2)`). Finally, *Bugula stolonifera* presents values that overlap both 0.67 and 0.75 at both temperatures (95\% CI at 10 ˚C: `r rounded(ci95Estimates[ci95Estimates$species == 'Stolonifera', 's10_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Stolonifera', 's10_97.5'], 2)`; 95\% CI at 25 ˚C: `r rounded(ci95Estimates[ci95Estimates$species == 'Stolonifera', 's25_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Stolonifera', 's25_97.5'], 2)`).

@@

That metabolic rate scales at an exponent of less than 1 across all four species (with the exception of *Hippopodina iririkiensis* measured at 10 ˚C) has interesting implications for our understanding of the energetics of growth for different colonial animals. For instance, evidence suggest that water clearance and feeding rates scale either isometrically or allometrically with colony size in sponges `r myCite(c('riisgard1993meps', 'kowalke2000jembe', 'mcmurray2014aquatbiol'))`, and these rates may follow the same mass scaling of metabolic rates (see e.g. `r myCite('thomassen1995meps')`). In contrast, most studies indicate that feeding rates scale isometrically or superlinearly in both encrusting and arborescent bryozoans (e.g. `r myCite(c('okamura1984jembe', 'okamura1985jembe', 'pratt2005meps'))`). Our results and those from other studies `r myCite('hartikainen2014jeb')` indicate that metabolic rates scale allometrically in arborescent 3D bryozoans. Consequently, while every unit increase in size results in proportional increase in feeding rate, every unit increase in size results in a less-than-proportional increase in metabolic rate. Thus, as colonies increase in size, their capacity to capture food increases more quickly than the rate at which they expend energy.

@@

In the wild, partial predation and physical disturbance are common sources of depression of metabolic rates in unitary organisms (e.g. `r myCite('glazier2011ecolmonogr')`) and partial mortality in colonial organisms `r myCite('sebens1987annurevecolsys')`. Our findings suggest that, for some colonial species, size reductions mediated by predation or disturbance might modulate the energetic demands of individuals of similar body mass in different environments, thus affecting life-history trajectories. For instance, at 25 ˚C, a 400-gram individual of *Bugula stolonifera* will spend ~`r round(metabolicDiff$metabolicDiff[metabolicDiff$species == 'Stolonifera'])`-fold more energy than a similar-sized individual at 10 ˚C. Conversely, due to the estimated positive interaction between mass-scaling and temperature, a similar-sized individual of *B. neritina* will spend ~`r round(metabolicDiff$metabolicDiff[metabolicDiff$species == 'Bugula'])`-fold more energy at 25 ˚C than at 10 ˚C. These differences, for instance, may help determine which species is a better competitor in different environments, as individual metabolic rates will also affect growth rates and time of reproduction onset `r myCite(c('sebens1987annurevecolsys', 'hart2009ecologyb'))`.

At the population level, disturbance and predation can also create different size distributions in colonial organisms. Our finding of allometric scaling of metabolic rate implies non-intuitive outcomes of such shifts in mean size of colonies in a population. For example, at 25 ˚C a population of 10 colonies of Microcionidae that each weigh 400 mg would have a cumulative metabolic demand of ~`r rounded(10 * exp(averageEstimates$lnB25[averageEstimates$species == 'Sponge'] + log(400)*averageEstimates$s25[averageEstimates$species == 'Sponge']), 0)` $\mu$l $\textrm{O}_\textrm{2}$ h^-1^. If a disturbance event halved each colony with minimal loss of biomass, then a population of those (now) 20 colonies (i.e. density twice as high), each weighing 200 mg, would have cumulative metabolic demand that is `r rounded((round(20 * exp(averageEstimates$lnB25[averageEstimates$species == 'Sponge'] + log(200)*averageEstimates$s25[averageEstimates$species == 'Sponge']))) / (round(10 * exp(averageEstimates$lnB25[averageEstimates$species == 'Sponge'] + log(400)*averageEstimates$s25[averageEstimates$species == 'Sponge']))), 1)`-fold higher (i.e. ~`r rounded(20 * exp(averageEstimates$lnB25[averageEstimates$species == 'Sponge'] + log(200)*averageEstimates$s25[averageEstimates$species == 'Sponge']), 0)` $\mu$l $\textrm{O}_\textrm{2}$ h^-1^), according to their new size. Thus, disturbance events change not only the size distribution of colonies but also increase the energy requirements of the population even if total standing biomass remains unchanged `r myCite('barneche2014ele')`. This new predation-mediated density dependence effect will increase total population energy flux, which is opposite to what is observed in unitary organisms that form groups or colonies, where group- or population-level metabolic rates often decrease due to increasing intraspecific competition (e.g. `r myCite('delong2014janimecol')`). We note, however, that a 'colony' here is a collection of physiologically connected interdependent zooids, which is different from a colony of independent unitary individuals, hence different mechanisms are likely to be operating as density of individuals changes.

@@

The dataset associated with this manuscript will be archived on DRYAD should the manuscript be accepted. All data and code necessary to reproduce this paper, its analyses, tables and figures can also be obtained on GitHub https://github.com/dbarneche/MTRBrEs.

@@

*Figure 2.* Estimates from the three-way interaction model as obtained by Bayesian methods in *JAGS*. Results for each species are represented in groups of four plots, with grey and black colors representing actual data and parameter estimates at 10 ˚C and 25 ˚C, respectively. *Top-left* and *bottom-right*: posterior probabilities of mass-scaling exponents and metabolic normalizations (respectively), with bars over curves representing 95\% credible intervals, and numbers representing average estimates. *Bottom-left*: covariance between mass-scaling exponents and metabolic normalizations for each of the 10,000 MCMC iterations (translucent circles), with overlaid density heat-map. *Top-right*: regressions of individual metabolic rate, $\textrm{ln}B_i$, as functions of individual body mass, $\textrm{ln}M_i$. Regression fits were plotted using the average from the posterior distributions. In probability posterior plots (*top-left* and *bottom-right*), smooth probability curves (*n* = 511 values) were approximated from posterior probability density curves (*n* = 512 values); differences between adjacent $x$ values (i.e. parameter estimates) were multiplied by the average of adjacent $y$ values (i.e. respective probability density values; $(x_{i+1} - x_{i}) \times ((y_{i+1} - y_{i})/2)$) thus insuring that the integral of the probability curve $\approx$ 1.

>>>>

*1. Animal collections*  
To collect our study organisms, roughened acetate sheets secured to the underside of 6 mm thick PVC sheets suspended from floating pontoons at a depth of 1 m were deployed at Manly Boat Harbour, Queensland, Australia (27˚27' S, 153˚11' E) and checked regularly to remove fouling organisms other than our species of interest: *Bugula neritina* and *B. stolonifera* (arborescent bryozoans that have a three dimensional tree-like growth form), *Hippopodina iririkiensis* (an encrusting bryozoan that grows mostly in two dimensions as flat disc), and an unidentified species of sponge belonging to the Microcionidae family (`r myCite('hart2009ecology')`; again a flat species that grows mostly in two dimensions) (Table 1). After 6 weeks in the field (average SST 22˚C), sheets bearing colonies were returned to the University of Queensland where they were kept in aerated seawater tanks (18˚C) for up to 48 h prior to measurements. For encrusting colonies, sections of acetate bearing whole colonies were then cut from the sheets and fragmented (see e.g. `r myCite(c('nakaya2005procb', 'white2011amnat'))` for details and tests of potential experimental artifacts). Arborescent colonies were cut directly from the acetate sheets and fragmented. We focused on non-reproductive colonies of bryozoans (at that time of year 6 weeks was not sufficient time for colonies to show reproduction) to avoid the potentially complicating effects that reproduction might bring. For the sponge, we found no evidence of reproduction, though we cannot rule out sperm production.

@@

*3. Respiration measurements*  
Individual metabolic rates (converted to $\mu$l $\textrm{O}_\textrm{2}$ h^-1^, *n* = `r nrow(numbersList$metRates)`), $B_i$, were measured as rate of oxygen consumption, $\dot{V}\textrm{O}_\textrm{2}$, of colony fragments at nominal temperatures of 10 or 25˚C. All individuals (regardless of size) and species were randomly assigned to test temperatures, and we used replicate vials and controlled temperature cabinets (see below). Measurements were conducted using a 144-channel PreSens Sensor Dish Reader (SDR, AS-1 Scientific, Wellington, New Zealand) according to standard techniques `r myCite(c('koster2008meps', 'lighton2008book', 'white2011amnat'))`.  $\dot{V}\textrm{O}_\textrm{2}$ was measured by placing the colony fragment in a glass vial containing 0.2$\mu$m filtered sea water and a non-consumptive $\textrm{O}_\textrm{2}$ sensor spot and calculated from the rate of change of $\textrm{O}_\textrm{2}$ saturation ($m_a$, \% h^‑1^) as:

@@

where $m_b$ is the rate of change of $\textrm{O}_\textrm{2}$ saturation for blank vials containing no animals (\% h^-1^), ${\beta}_{\textrm{O}_\textrm{2}}$ is the oxygen capacitance of air-saturated sea water at 10˚C (6.4 ml l^-1^) or 25 ˚C (4.74 ml l^-1^) `r myCite('cameron1986book')` and $V$ is water volume (vials were 5 ml in volume and water volume was calculated by subtracting the volume of acetate and animals). At least four blank vials were recorded simultaneously with each 24-well plate to account for microbial oxygen consumption and sensor spots were calibrated with air-saturated seawater (100\% A.S.) and water containing 2\% sodium sulfite (0\% A.S.). After a 1 h settling-in period (i.e. time necessary for the system to come into equilibrium), $\dot{V}\textrm{O}_\textrm{2}$ measurements lasted up to 20 h depending on individual mass (tiny individuals require much longer measurement periods in order to detect $\dot{V}\textrm{O}_\textrm{2}$). Following measurement of $\dot{V}\textrm{O}_\textrm{2}$, colony fragments were blotted dry and weighed to 0.1 mg (Sartorius A 200 S, Sartorius AG, Göttingen, Germany). All measurements of $\dot{V}\textrm{O}_\textrm{2}$ were made in one of four dark constant temperature cabinets, with temperatures within a run randomized among cabinets.

@@

where mass, $\textrm{ln}M_i$ (mg), is a continuous variable, while temperature (two levels: 10˚C and 25˚C) and species (four levels corresponding to the four species) are categorical variables. With this model, we allow each species to have a different mass-scaling slope that may change with temperature; we also account for species- and temperature-specific deviations on the metabolic normalization (i.e. model intercept). We also included date of experiment, $\Delta D$ as a random effect on the metabolic normalization to help control for natural experimental variability. We then performed a model selection procedure (see below) by comparing the above full model with a series of nested two- and one-way interaction models in order to formally test for mass-temperature interactions in metabolic rates among different species.

All models were initially fitted using maximum likelihood `r myCite('zuur2009book')` with the R package *lme4* version 1.1-8 `r myCite('lme4')`. Each nested model was compared against the full model using likelihood ratio tests, and models were considered significantly different if *p* < 0.05. We then fitted the best model in a Bayesian framework by calling *JAGS* version 3.4.0 from the R package *R2jags* version 0.05-03 `r myCite('r2jags')` in order to derive posterior distributions and associated 95\% credible intervals (CIs) for the fitted parameters. Random effects were assumed to be normally distributed, with means of 0. Fitted parameters were assigned priors that were vague (i.e. locally uniform over the region supported by the likelihood) `r myCite('kruschke2014book')`. The posterior distributions of model parameters were estimated using Markov chain Monte Carlo (MCMC) methods by constructing three chains of 500,000 steps each, including 250,000-step burn-in periods. Chains were then updated with 500,000 steps each until convergence was reached using 250-thinning step interval, so a total of 6,000 steps were retained to estimate posterior distributions (i.e. 3 $\times$ 500,000/250 = 6,000). All data and code necessary to reproduce this paper, its analyses, tables and figures can be obtained on GitHub https://github.com/dbarneche/MTRBrEs `r myCite('barneche2016ecologyData')`.

@@

The model selection indicated that the three-way interaction model (equation 2) performed significantly better than all nested two- and one-way interaction models (Table 2), thus implying that different species present different interactions between mass-scaling and temperature dependence. Particularly, the best model yields no interaction between mass scaling and temperature dependence for *Bugula stolonifera* and the Microcionidae sponge (large overlap between 95\% credible intervals for the mass-scaling slopes, $\alpha$) (Fig. 2). Still, while the mass-scaling slopes are statistically the same between temperatures for these species, there is evidence for higher metabolic normalizations (i.e. intercepts), $B_o$, at higher temperatures (though marginally significant for Microcionidae, i.e. slight overlap between 95\% credible intervals; bottom-right quadrants in Fig. 2). For example, metabolic rates increase `r rounded(exp(averageEstimates$lnB25[averageEstimates$species == 'Stolonifera'])/exp(averageEstimates$lnB10[averageEstimates$species == 'Stolonifera']))`-fold between 10˚C and 25˚C for *B. stolonifera*, yielding a $Q_{10}$ of `r rounded(q10andErs$q10[q10andErs$species == 'Stolonifera'])`, which is equivalent to an activation energy of `r rounded(q10andErs$er[q10andErs$species == 'Stolonifera'], 2)` eV `r myCite(c('gillooly2001','yvondurocher2012a'))`. The best model indicates that the difference in metabolic rates between 25˚C and 10˚C for *Hippopodina iririkiensis* decreases, on average, `r rounded(tempDeltas[1,'tempDelta'])`-fold moving from `r round(exp(2))` to `r round(exp(5))` mg. For *Bugula neritina*, however, rates increase `r rounded(1/tempDeltas[3,'tempDelta'])`-fold along the same body-mass gradient.

Average posterior estimates of mass-scaling exponents (top-left quadrants in Fig. 2) are inconsistent among the studied species and between growth forms (2D and 3D). For example, for organisms that exhibit 2D growth, *Hippopodina iririkiensis* shows exponents that encompass both 0.67 (Fractal geometry model prediction for 2D organisms; `r myCite('koontz2009amjbot')`) and 0.75 (Fractal geometry model prediction for 3D organisms; `r myCite(c('west1997science', 'west1999science'))`) at 10˚C (95\% CI: `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's10_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's10_97.5'], 2)`), but that are significantly shallower than 0.75 at 25˚C (95\% CI: `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's25_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's25_97.5'], 2)`). Microcionidae presents exponents that only overlap 0.67 at 10˚C, but not at 25˚C (95\% CI at 10˚C: `r rounded(ci95Estimates[ci95Estimates$species == 'Sponge', 's10_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Sponge', 's10_97.5'], 2)`; 95\% CI at 25˚C: `r rounded(ci95Estimates[ci95Estimates$species == 'Sponge', 's25_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Sponge', 's25_97.5'], 2)`). On the other hand, for 3D organisms, *Bugula neritina* presents scaling exponents that overlap both 0.67 and 0.75 at 25˚C (95\% CI: `r rounded(ci95Estimates[ci95Estimates$species == 'Bugula', 's25_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Bugula', 's25_97.5'], 2)`), but are significantly shallower at 10˚C (95\% CI: `r rounded(ci95Estimates[ci95Estimates$species == 'Bugula', 's10_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Bugula', 's10_97.5'], 2)`). Finally, *Bugula stolonifera* presents values that overlap both 0.67 and 0.75 at both temperatures (95\% CI at 10˚C: `r rounded(ci95Estimates[ci95Estimates$species == 'Stolonifera', 's10_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Stolonifera', 's10_97.5'], 2)`; 95\% CI at 25˚C: `r rounded(ci95Estimates[ci95Estimates$species == 'Stolonifera', 's25_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Stolonifera', 's25_97.5'], 2)`).

@@

That metabolic rate scales at an exponent of less than 1 across all four species (with the exception of *Hippopodina iririkiensis* measured at 10˚C) has interesting implications for our understanding of the energetics of growth for different colonial animals. For instance, evidence suggest that water clearance and feeding rates scale either isometrically or allometrically with colony size in sponges `r myCite(c('riisgard1993meps', 'kowalke2000jembe', 'mcmurray2014aquatbiol'))`, and these rates may follow the same mass scaling of metabolic rates (see e.g. `r myCite('thomassen1995meps')`). In contrast, most studies indicate that feeding rates scale isometrically or superlinearly in both encrusting and arborescent bryozoans (e.g. `r myCite(c('okamura1984jembe', 'okamura1985jembe', 'pratt2005meps'))`). Our results and those from other studies `r myCite('hartikainen2014jeb')` indicate that metabolic rates scale allometrically in arborescent 3D bryozoans. Consequently, while every unit increase in size results in proportional increase in feeding rate, every unit increase in size results in a less-than-proportional increase in metabolic rate. Thus, as colonies increase in size, their capacity to capture food increases more quickly than the rate at which they expend energy.

@@

In the wild, partial predation and physical disturbance are common sources of depression of metabolic rates in unitary organisms (e.g. `r myCite('glazier2011ecolmonogr')`) and partial mortality in colonial organisms `r myCite('sebens1987annurevecolsys')`. Our findings suggest that, for some colonial species, size reductions mediated by predation or disturbance might modulate the energetic demands of individuals of similar body mass in different environments, thus affecting life-history trajectories. For instance, at 25˚C, a 400-gram individual of *Bugula stolonifera* will spend ~`r round(metabolicDiff$metabolicDiff[metabolicDiff$species == 'Stolonifera'])`-fold more energy than a similar-sized individual at 10˚C. Conversely, due to the estimated positive interaction between mass-scaling and temperature, a similar-sized individual of *B. neritina* will spend ~`r round(metabolicDiff$metabolicDiff[metabolicDiff$species == 'Bugula'])`-fold more energy at 25˚C than at 10˚C. These differences, for instance, may help determine which species is a better competitor in different environments, as individual metabolic rates will also affect growth rates and time of reproduction onset `r myCite(c('sebens1987annurevecolsys', 'hart2009ecologyb'))`.

At the population level, disturbance and predation can also create different size distributions in colonial organisms. Our finding of allometric scaling of metabolic rate implies non-intuitive outcomes of such shifts in mean size of colonies in a population. For example, at 25˚C a population of 10 colonies of Microcionidae that each weigh 400 mg would have a cumulative metabolic demand of ~`r rounded(10 * exp(averageEstimates$lnB25[averageEstimates$species == 'Sponge'] + log(400)*averageEstimates$s25[averageEstimates$species == 'Sponge']), 0)` $\mu$l $\textrm{O}_\textrm{2}$ h^-1^. If a disturbance event halved each colony with minimal loss of biomass, then a population of those (now) 20 colonies (i.e. density twice as high), each weighing 200 mg, would have cumulative metabolic demand that is `r rounded((round(20 * exp(averageEstimates$lnB25[averageEstimates$species == 'Sponge'] + log(200)*averageEstimates$s25[averageEstimates$species == 'Sponge']))) / (round(10 * exp(averageEstimates$lnB25[averageEstimates$species == 'Sponge'] + log(400)*averageEstimates$s25[averageEstimates$species == 'Sponge']))), 1)`-fold higher (i.e. ~`r rounded(20 * exp(averageEstimates$lnB25[averageEstimates$species == 'Sponge'] + log(200)*averageEstimates$s25[averageEstimates$species == 'Sponge']), 0)` $\mu$l $\textrm{O}_\textrm{2}$ h^-1^), according to their new size. Thus, disturbance events change not only the size distribution of colonies but also increase the energy requirements of the population even if total standing biomass remains unchanged `r myCite('barneche2014ele')`. This new predation-mediated density dependence effect will increase total population energy flux, which is opposite to what is observed in unitary organisms that form groups or colonies, where group- or population-level metabolic rates often decrease due to increasing intraspecific competition (e.g. `r myCite('delong2014janimecol')`). We note, however, that a 'colony' here is a collection of physiologically connected interdependent zooids, which is different from a colony of independent unitary individuals, hence different mechanisms are likely to be operating as density of individuals changes.

@@

All data and code necessary to reproduce this paper, its analyses, tables and figures can be obtained on GitHub https://github.com/dbarneche/MTRBrEs (`r myCite('barneche2016ecologyData')`, doi: 10.5281/zenodo.159736).

@@

*Figure 2.* Estimates from the three-way interaction model as obtained by Bayesian methods in *JAGS*. Results for each species are represented in groups of four plots, with grey and black colors representing actual data and parameter estimates at 10˚C and 25˚C, respectively. *Top-left* and *bottom-right*: posterior probabilities of mass-scaling exponents and metabolic normalizations (respectively), with bars over curves representing 95\% credible intervals, and numbers representing average estimates. *Bottom-left*: covariance between mass-scaling exponents and metabolic normalizations for each of the 10,000 MCMC iterations (translucent circles), with overlaid density heat-map. *Top-right*: regressions of individual metabolic rate, $\textrm{ln}B_i$, as functions of individual body mass, $\textrm{ln}M_i$. Regression fits were plotted using the average from the posterior distributions. In probability posterior plots (*top-left* and *bottom-right*), smooth probability curves (*n* = 511 values) were approximated from posterior probability density curves (*n* = 512 values); differences between adjacent $x$ values (i.e. parameter estimates) were multiplied by the average of adjacent $y$ values (i.e. respective probability density values; $(x_{i+1} - x_{i}) \times ((y_{i+1} - y_{i})/2)$) thus insuring that the integral of the probability curve $\approx$ 1.
